An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2021-07-27 latex_files/ch2/2_1_LFI_intro.tex
minor improvements

An implicit or simulator-based model is a parameterised stochastic
data generating mechanism \cite{Gutmann2016}. The key characteristic
of these models is that although we can sample (simulate) data points,
we cannot evaluate the likelihood of a specific set of observations
$\data$. Formally, a simulator-based model is described as a
parameterised family of probability density functions
$\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$, whose closed-form is either
unknown or intractable to evaluate. Whereas evaluating
$p_{\yb|\thetab}(\yb)$ is intractable, sampling is
feasible. Practically, a simulator can be understood as a black-box
machine $M_r$\footnote{The subscript $r$ in $M_r$ indicates the
  \textit{random} simulator. In the next chapters we will introduce
  $M_d$ witch stands for the \textit{deterministic} simulator.} that
given a set of parameters $\thetab$, produces samples $\yb$ in a
stochastic manner i.e.\ $M_r(\thetab) \rightarrow \yb$.

Simulator-based models are particularly captivating due to the
modelling freedom they provide; any physical process that can be
conceptualised as a computer program of finite steps can be modelled
as a simulator-based model without any compromise. The modelling
freedom includes any amount of hidden (unobserved) internal variables
or rule-based decisions. Hence, implicit models are often used to
model physical phenomena in many fiels such as biology, neuroscience
etc. This degree of freedom comes at a cost; performing the inference
is particularly demanding from both the computational and the
mathematical perspective. The class of methods that try to approximate
the posterior distribution in these cases are called likelihood-free
inference methods. Unfortunately, the algorithms deployed so far
permit the inference only at low-dimensional parametric spaces, i.e.\
$\thetab \in \mathbb{R}^D$ where $D$ is small\footnote{In broad terms,
  small refers to dimensionality $D$ less than a hundred.}.

@@

As in most Machine Learning (ML) concepts, the fundamental goal is the
derivation of one (or many) parameter configuration(s) $\thetab^*$
that \textit{describe} the data best i.e.\ generate samples
$M_r(\thetab^*)$ that are as close as possible to the observed data
$\data$. Following the Bayesian approach, we treat the
parameters of interest $\thetab$ as random variables and we try to
\textit{infer} a posterior distribution $p(\thetab|\data)$ on them.

>>>>

An implicit or simulator-based\footnote{The terms can be used
  interchangeably} model is a parameterised stochastic data generating
mechanism \cite{Gutmann2016}. The key characteristic of these models
is that although it is feasible to sample (simulate) data points, it
is impossible to evaluate the likelihood. Formally, a simulator-based
model is a parameterised family of probability density functions
$\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$ whose closed-form is either
unknown or intractable to evaluate. Practically, a simulator can be
understood as a black-box machine $M_r$\footnote{The subscript $r$ in
  $M_r$ indicates the \textit{random} simulator. In the next chapters
  we will introduce $M_d$ witch stands for the \textit{deterministic}
  simulator.} that given a set of parameters $\thetab$, produces
samples $\yb$ in a stochastic manner i.e.\
$M_r(\thetab) \rightarrow \yb$. Simulator-based models are
particularly captivating due to the modelling freedom they provide;
any physical process, that can be conceptualised as a computer program
of finite steps, can be represented as a simulator-based model. For
this reason, they are quite popular in many fiels such as biology,
neuroscience etc. This degree of freedom comes at a cost; performing
the inference is particularly demanding from both computational and
mathematical perspective.

@@

As in most Machine Learning (ML) concepts, the fundamental goal is the
derivation of one (or many) parameter configuration(s) $\thetab^*$
that \textit{describe} the data best i.e.\ generate samples
$M_r(\thetab^*)$ that are as close as possible to the observed data
$\data$. Following the Bayesian approach, we treat the parameters of
interest $\thetab$ as random variables and we try to \textit{infer} a
posterior distribution $p(\thetab|\data)$ on them. The class of
methods for approximate the posterior distribution in implicit models
is called likelihood-free inference methods. Unfortunately, the
algorithms deployed so far permit the inference only at
low-dimensional parametric spaces, i.e.\ $\thetab \in \mathbb{R}^D$
where $D$ is small\footnote{In broad terms, small refers to
  dimensionality $D$ less than a hundred.}.

### 2021-07-27 latex_files/ch2/2_3_OMC.tex
minor improvements

For describing the likelihood approximation introduced in OMC, it is
convenient to firstly define the indicator function based on
$\region(\yb)$. The indicator function $\indicator{\region(\yb)}(\xb)$
returns 1 if $\xb \in \region(\yb)$ and 0 otherwise. If
$d(\cdot,\cdot)$ is a formal distance, due to symmetry
$\indicator{\region(\yb)}(\xb) = \indicator{\region(\xb)}(\yb)$, so
the expressions can be used interchangeably.

>>>>

For describing the likelihood approximation introduced in OMC, it is
convenient to firstly define the indicator function
$\indicator{\region(\yb)}(\xb)$, that returns 1 if
$\xb \in \region(\yb)$ and 0 otherwise. If $d(\cdot,\cdot)$ is a
formal distance, due to symmetry
$\indicator{\region(\yb)}(\xb) = \indicator{\region(\xb)}(\yb)$, so
the expressions can be used interchangeably.

### 2021-07-27 latex_files/ch3/3_2_2_training.tex
fixed warnings
minor improvements

\begin{Code}
------------------------------ python snippet ------------------------------
  # Trainining (fitting) part
  n1 = 500 # number of optimisation problems
  seed = 21 # seed for solving the optimisation problems
  eps = .75 
  use_bo = False # set to True for switching to Bayesian optimisation

@@

In figure \ref{fig:example_training_hist} we illustrate the
distribution of the distances obtained and the acceptance area of the
first optimisation problem. We observe that most optimal points
produce almost zero distance.

@@

\begin{figure}[h]
  \begin{center}    
    \input{latex_files/images/chapter3/example_gt.tex}
    \input{latex_files/images/chapter3/example_posterior.tex}
  \end{center}
  \caption[Histogram of distances at the 1D example.]{Histogram of
    distances and visualisation of a specific region.}
  \label{fig:example_training_hist}
\end{figure}

>>>>

\begin{Code}
------------------------------ python snippet ------------------------------
  # Training (fitting) part
  n1 = 500 # number of optimisation problems
  seed = 21 # seed for solving the optimisation problems
  eps = .75 
  use_bo = False # set to True for switching to Bayesian optimisation

@@

@@

\begin{figure}[ht]
  \begin{center}    
    \input{latex_files/images/chapter3/example_gt.tex}
    \input{latex_files/images/chapter3/example_posterior.tex}
  \end{center}
  \caption[Histogram of distances at the 1D example.]{Histogram of
    distances and visualisation of a specific region.}
  \label{fig:running_example_romc_inference}
\end{figure}
